---
output:
  md_document:
    variant: markdown_github
---


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Travis build status](https://travis-ci.org/talegari/bigdist.svg?branch=master)](https://travis-ci.org/talegari/bigdist)

# bigdist

bigdist package facilitates storing distance matrices on disk as 'file-backed matrix' (FBM) using [bigstatsr](https://cran.r-project.org/package=bigstatsr) package. The FBM stores a symmetric matrix. Each distance is stored as a 'float/double' approximately requiring 4 or 8 bytes. The resulting file size will approximately be 4 or 8 * size^2 where size is nrow/ncol of the data matrix. Working with bigdist object requires working knowledge of bigstatsr package.

[disto](https://cran.r-project.org/package=disto) package provides a unified interface to distance matrices in-memory (class 'dist') or on disk (class 'bigdist').

## Installation

You can install bigdist from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("talegari/bigdist")
```

## Functionality

```{r}
library("bigdist")

# create a random matrix
set.seed(1)
amat <- matrix(rnorm(1e4), ncol = 10)
td   <- tempdir()
dir.create(td)

# create a bigdist object with FBM(fle-baked matrix) on disk
temp <- bigdist(mat = amat, file = file.path(td, "tempfile"))
temp
temp$fbm$backingfile
temp$fbm[1, 2]

# connect to FBM on disk as a bigdist object
temp2 <- bigdist(file = file.path(td, "tempfile_1000_float"))
temp2
temp2$fbm[1,2]

# check the size of bigdist object
bigdist_size(temp)

# bigdist accessors

# ij
bigdist_extract(temp, 1, 2)
bigdist_extract(temp, 1:2, 3:4)
bigdist_extract(temp, 1:2, 3:4, product = "inner")
dim(bigdist_extract(temp, 1:2, ))
dim(bigdist_extract(temp, , 3:4))

# k (lower trianle indexing)
bigdist_extract(temp, k = 3:7)

# bigdist replacers

# ij
bigdist_replace(temp, 1, 2, 10)
bigdist_extract(temp, 1, 2)
bigdist_replace(temp, 1:2, 3:4, 11:12)
bigdist_extract(temp, 1:2, 3:4, product = "inner")

# k (lower trianle indexing)
bigdist_replace(temp, k = 3:7, value = 51:55)
bigdist_extract(temp, k = 3:7)

# subset a bigdist object
temp_subset <- bigdist_subset(temp, index = 201:300, file = file.path(td, "tempfile2"))
temp_subset
temp_subset$fbm$backingfile

# convert a dist object(in memory) to a bigdist object
temp3 <- as_bigdist(dist(mtcars), file = file.path(td, "tempfile3"))
temp3

# remove the FBM from disk
unlink(td, recursive = TRUE)
```

